---
title: "project_MAP_569"
author: "Shiqi liu"
date: "20/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE)
library(readr)
library(GGally)
library(dplyr)
library(ggfortify)
library(corrplot)
library(car)



```

## R Markdown
```{r}
abalone <- read_csv("Abalone_data.csv")
abalone$Sex <- as.factor(abalone$Sex)
abalone = abalone[is.finite(log(abalone$Height)), ]


# train test split
train_test_spit <- function(data, train_percent){
  smp_size <- floor(train_percent * nrow(data) /100)
  train_ind <- sample(seq_len(nrow(data)), size = smp_size)
  set.seed(1)
  train <- data[train_ind, ]
  test <- data[-train_ind, ]
  list(train = train, test = test)
}

abalone_split <- train_test_spit(abalone,80)
train <- abalone_split$train
test <- abalone_split$test



## 75% of the sample size
```


```{r}
ggpairs(abalone, lower = list(continuous = wrap(ggally_points, size = 0.1, mapping = ggplot2::aes(colour=Sex))), columns = 2:9,
        ) 

r=round(cor(abalone[,2:9]),2); corrplot(r,method="ellipse")




```

```{r}
# plot(abalone$Whole_weight,log(abalone$Rings))
# plot(abalone$Whole_weight,abalone$Rings)
# plot(log(abalone$Whole_weight),log(abalone$Rings))
# plot(abalone$Height,log(abalone$Rings))
# plot(log(abalone$Height),log(abalone$Rings))


# plot(log(abalone$Diameter),log(abalone$Rings))

feature_engineering <- function(abalone_raw, ancova = FALSE){
  abalone_transform <- abalone_raw[-c(2:3,4:9)]
  abalone_transform$log_Length = log(abalone_raw$Length)
  abalone_transform$log_Diameter = log(abalone_raw$Diameter)
  abalone_transform$log_Height = log(abalone_raw$Height)
  abalone_transform$log_Shucked_weight = log(abalone_raw$Shucked_weight) 
  abalone_transform$log_Viscera_weight = log(abalone_raw$Viscera_weight) 
  abalone_transform$log_Shell_weight = log(abalone_raw$Shell_weight) 
  abalone_transform$log_Whole_weight = log(abalone_raw$Whole_weight) 
  abalone_transform$sqrt_log_Ring = sqrt(log(abalone_raw$Rings))
  # abalone_transform = abalone_transform %>% filter(is.finite(abalone_transform$log_Height))
  # abalone_transform$log_Height[!is.finite(abalone_transform$log_Height)] <- 3.3
  
  if(ancova){
    abalone_transform$Sex[abalone_transform$Sex == 'M'] <- 'F'
  }
  
  
  abalone_transform
} 


abalone_transformed = feature_engineering(abalone)

# ggpairs(abalone_transformed, upper = list(continuous = wrap("cor", family="sans")),lower = list(continuous = wrap(ggally_points, size = 0.1, mapping = ggplot2::aes(colour=Sex))), columns = 2:9,
#         )
```


```{r}


abalone_splited = train_test_spit(abalone,80)

train = abalone_splited$train
test = abalone_splited$test
train_transformed = feature_engineering(train)
test_transformed = feature_engineering(test)

```



```{r}
data_1 <- train_transformed%>% filter(log_Height<4, log_Height>2.5)
plot(train_transformed$log_Height,train_transformed$sqrt_log_Ring)

model_height <- data_1%>% lm(formula = sqrt_log_Ring ~ log_Height)
plot(model_height)
ncvTest(model_height)

summary(model_height)

```

## Question 8
```{r}

ggpairs(abalone_transformed, upper = list(continuous = wrap("cor", family="sans")),lower = list(continuous = wrap(ggally_points, size = 0.1, mapping = ggplot2::aes(colour=Sex))), columns = 2:9,
        ) 


```
Some remarks:
* some features are strongly correlated, is it a problem?
* for the moment, we shall use log_Whole_weight, log_Shell_weight, log_Height, and log_diameter which has the largest correlation with the log_Ring
* explain the form of the plot
```{r}


model_1 <-  data_1 %>% lm(formula = sqrt_log_Ring ~  log_Height + log_Shell_weight + log_Whole_weight + log_Diameter)

summary(model_1)

plot(model_1)

# Breush-Pagan test the reason why choose model_1
ncvTest(model_1)



# Shapiro-Wilk test
# shapiro.test(residuals(model_0)) 
# shapiro.test(residuals(model_1)) 

# remove the point 236 with a cook distance larger than 1, then refit the model
# data_0 <- data_0[-c(236),]
# model_0 <-  data_0 %>% lm(formula = log_Ring ~ log_Height + log_Shell_weight + log_Whole_weight + log_Diameter)
# summary(model_0)
# plot(model_0)



```
## Question 9

As we can see from the result below, ANOVA suggests that our multilinear model is better than simple linear model
```{r}
anova(model_height,model_1)

```


```{r}
# anvoca
abalone %>% ggplot() + aes(x = Sex, y = Rings, color = Sex) + geom_boxplot(alpha=0.5,) + geom_jitter(width = 0.25)
model_sex= data_1 %>% lm(formula = sqrt_log_Ring ~   Sex)
summary(model_sex)




data_2 = data.frame(data_1)

# explain why doing this
model_ancova = data_2 %>% lm(formula = sqrt_log_Ring ~  log_Height + log_Shell_weight*Sex + log_Whole_weight*Sex + log_Diameter)
summary(model_ancova)
data_2$Sex[data_1$Sex == 'M'] <- 'F'
model_ancova = data_2 %>% lm(formula = sqrt_log_Ring ~  log_Height + log_Shell_weight*Sex + log_Whole_weight*Sex + log_Diameter)
summary(model_ancova)

# Question???
```
## performance on trainning set

```{r}
train_error <- function(lm) mean(lm$residuals^2)

round_to_integer <- function(number){
  if (number - floor(number) > 0.5) {
    return(ceiling(number))
  }
  return(floor(number))
}

back_to_Ring <- function(vector){
  vector_inversed = exp(vector^2)
  as.double(lapply(vector_inversed,round_to_integer))
}

loss <- function(y, y_predict){
  # return(mean((y -y_predict)^2))
  return(mean(abs(y -y_predict)))
}

predict_as_ring <- function(lm, X_test, ancova = FALSE){
  return( back_to_Ring(predict(lm, feature_engineering(X_test,ancova = ancova))))
}

predict_error <- function(lm,X_test,ancova = FALSE) {
  return(loss(X_test$Rings, predict_as_ring(lm, X_test, ancova = ancova) ) )
}

```

```{r}
predict_error(model_height, test)
predict_error(model_1, test)
predict_error(model_ancova, test,ancova = TRUE)

```
```{r}
# let me see


result = data.frame(y = test$Rings, y_predict = predict_as_ring(model_ancova,test,ancova = TRUE))

```



